Fast and adaptive cointegration based model for forecasting high frequency financial time series

### 2016-04-25 articles-conferences/iscef2016/src/intro.tex
agregando imágenes nuevas

VECM parameters are obtained using the ordinary least squares
(OLS) method~\cite{golub1980}. Since OLS involves many calculations, the parameter estimation is computationally expensive when the number of lagged values and
data increases. Moreover, obtaining cointegration vectors 
is also an expensive routine. This is a main limitation to use VECM with stream data with high frequency.
Chen and Lung's BVECM \cite{chen2003} addresses the advantage of distributed processing over conventional rolling window processing.

Our aims were to study if a parallel version of VECM can be used with high frequency stream data.
Our approach was to determine adaptively the number of observations and lags of VECM which
maximise cointegration relations in the past. 
 This search is computationally
expensive because the Johansen
method~\cite{johansen1995} is required to find cointegration vectors, which is of order $O(n^3)$. Therefore, our proposal, called Adaptively Vector Error Correction (AVECM), 
is to parallelise this search in order to get new parameters before new data arrives.
Model effectiveness is focused on out-of-sample forecast rather than in-sample
fitting. This criterion allows AVECM prediction capability to be
expressed rather than just explaining data history.
The forecast capability of our method was measured using MSE and the Theil's $U$-statistic \cite{theil1966} widely used in economic forecast. Tests were run using four currency rates: Euro (EUR) to
United States Dollar (USD) (EURUSD), British Pound (GBP) to USD (GBPUSD), USD to Swiss
Franc (CHF) (USDCHF) and USD to Japanese Yen (JPY) (USDJPY) with a 10-seconds frequency.

This paper is organised as follows: section~\ref{sec:background} presents the
VAR and VECM, the AVECM algorithm is presented in section~\ref{sec:methodology}.
In section~\ref{sec:results} we describe the tests carried on to assess the
accuracy and the execution time of AVECM.  This section also includes a
description of the test data.  Section~\ref{sec:conclusions} contains the
conclusions and a discussion of future research.

>>>>

VECM parameters are obtained using the ordinary least squares
(OLS) method \cite{golub1980}. Since OLS involves many calculations, the parameter estimation is computationally expensive when the number of lagged values and
data increases. Moreover, obtaining cointegration vectors 
is also an expensive routine. This is a main limitation to use VECM with stream data with high frequency.
Chen and Lung's BVECM \cite{chen2003} addresses the advantage of distributed processing over conventional rolling window processing.

Our aims were to study if a parallel version of VECM can be used with high frequency stream data.
Our approach was to determine adaptively the number of observations and lags of VECM which
maximise cointegration relations in the past. 
 This search is computationally
expensive because the Johansen
method \cite{johansen1995} is required to find cointegration vectors, which is of order $O(n^3)$. Therefore, our proposal, called Adaptively Vector Error Correction (AVECM), 
is to parallelise this search in order to get new parameters before new data arrives.
Model effectiveness is focused on out-of-sample forecast rather than in-sample
fitting. This criterion allows AVECM prediction capability to be
expressed rather than just explaining data history.
The forecast capability of our method was measured using MSE and the Theil's $U$-statistic \cite{theil1966} widely used in economic forecast. Tests were run using four currency rates: Euro (EUR) to
United States Dollar (USD) (EURUSD), British Pound (GBP) to USD (GBPUSD), USD to Swiss
Franc (CHF) (USDCHF) and USD to Japanese Yen (JPY) (USDJPY) with a 10-seconds frequency.

This paper is organised as follows: section \ref{sec:background} presents the
VAR and VECM, the AVECM algorithm is presented in section \ref{sec:methodology}.
In section \ref{sec:results} we describe the tests carried on to assess the
accuracy and the execution time of AVECM.  This section also includes a
description of the test data.  Section \ref{sec:conclusions} contains the
conclusions and a discussion of future research.

### 2016-04-25 articles-conferences/iscef2016/src/methodology.tex
agregando imágenes nuevas

The same experiments, this time using $p=3$, were carried out in order to see the
effect on the number of cointegration vectors found. 
Figure~\ref{fig:histsp3} shows that $p$ affects the distribution of the number of cointegration vectors by small amounts.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.9\textwidth]{img/histCointVectors-offset20520-p-3-freq-10s}
  \caption{Histogram of the number of cointegration vectors using $p=3$. The distribution is very similar to the one with $p=1$}
   \label{fig:histsp3}
\end{figure}

@@

\begin{figure}[ht!]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/MSE-offset20520-p-2-freq-10s}
  \caption{MSE versus the percentage of cointegration considering 1000
  iterations. Best $L$ found was 1600. Below MSE versus $L$ shows a rapidly decreasing behaviour, founding minimum at PC = 84.3\% }
  \label{fig:cointvsmse}
\end{figure}

>>>>

@@

\begin{figure}[ht!]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/MSE-offset20520-p-1-freq-10s}
  \caption{MSE versus the percentage of cointegration considering 1000
  iterations. Best $L$ found was 1600. Below MSE versus $L$ shows a rapidly decreasing behaviour, founding minimum at PC = 84.3\% }
  \label{fig:cointvsmse}
\end{figure}
